---
title: "Using the gt-package and rayshader with the ChainLadder-package"
author: "Andreas Slåttelid"
output: 
    word_document:
        toc: yes
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
The idea here is to show, how the gt-package could be used in combination with the ChainLadder-package in R. What makes this difficult is the transformation from the socalled "triangle"-object in ChainLadder, to a Tidyverse-dataframe.

Furthermore, we will also see how one could plot the chainladder triangle in 3d.

\newpage
# ChainLadder
```{r libraries}
library(tidyverse)    # wrangeling of data + %>%
library(gt)           # Creating runoff triangles
library(ChainLadder)  # contains the RAA dataset
library(rayshader)    # 3d-plot of triangle
library(viridis)      # coloring of heatmap
library(viridisLite)  # coloring of heatmap
```  

In the [ChainLadder-package](https://cran.r-project.org/web/packages/ChainLadder/vignettes/ChainLadder.html), there is a data set called RAA, the goal will be to go from:

```{r}
triangle <- RAA
triangle
```

And end up with the table displayed on the **last page** :)


The gt-package is all about formatting tables nicely, and it works very good with dataframes, however, we have:
```{r}
class(RAA)
```
namely a triangle/matrix object, the job will therefore be to transform objects from triangles to dataframes, which is not as obvious as one should expect

\newpage
# Transformation process

```{r triangle_to_matrix}
triangle_tibble <- triangle %>% as_tibble() %>% cbind(rownames(triangle), .)

#corresponds to development years j: 1,2, ...
development_years <- colnames(triangle_tibble)[2:length(colnames(triangle_tibble))]

colnames(triangle_tibble) <- c("Accident year i", development_years)

#select everything except "Accident year"
tmp <- triangle_tibble %>% 
  select(-"Accident year i") %>% 
  as.character()

#matrix of zeros
m <- matrix(0, ncol = length(development_years), 
               nrow = length(rownames(triangle)))

#append the data from traingle object to matrix
for (i in 1:length(development_years)){
  m[i, ] <- eval(parse(text = tmp[i])) #evaluate vector camouflaged as character.
}
m
```
```{r matrix_to_df}
#store the matrix as a dataframe
df <- data.frame(m)
colnames(df)
#include the accident years:
df <- df %>% 
  mutate(year = rownames(RAA), .before = X1)

colnames(df) <- c("year", development_years)
```

\newpage 
# Display ChainLadder as gt-table

```{r gt}
maximum_color_value <- max(m, na.rm = TRUE)

columns_formatted <- colnames(df)[2:length(colnames(df))]

gt_triangle <- df %>% 
  gt(rowname_col = "year") %>% 
  tab_stubhead(label = "Accident year i") %>% 
  fmt_number(columns = columns_formatted, 
             decimals = 0,
             accounting = TRUE)  %>% 
  tab_spanner(
    label = "Development year j",
    columns = all_of(colnames(df))
  ) %>% 
  sub_missing(missing_text = "-") %>% 
  data_color(
    columns = columns_formatted,
    colors = scales::col_numeric(
      palette = c("blue", "green", "orange", "red"),
      domain = c(0, maximum_color_value)
    )
  )

gt_triangle
```


\newpage 
# 3D-plot of ChainLadder 

```{r}
#replace NA's with 0
df2 <- data.frame(m)
df2[is.na(df2)] <- 0

df2 <- df2 %>% 
  mutate(year = rownames(RAA), .before = X1)

colnames(df2) <- c("year", colnames(RAA))
```




```{r wide_to_long_df}
df2_long <- df2 %>%
  pivot_longer(!year, names_to = "development_year", values_to = "claims") %>% 
  mutate(development_year_helper = as.numeric(development_year))

#sort properly development year, had 1, 10, 2, 3, ..., 9
df2_long$development_year <-  reorder(df2_long$development_year, df2_long$development_year_helper)

triangle_dim2_plt <- ggplot(df2_long, aes(year, development_year, fill= claims)) +
  geom_tile() + 
  scale_fill_viridis()
triangle_dim2_plt
```

\newpage 

Using the [rayshader-package](https://www.rayshader.com/reference/plot_gg.html#examples), one can easily go from a 2d-ggplot to a 3d-plot. 

```{r}
#triangle_dim3_plt <- rayshader::plot_gg(triangle_dim2_plt,
#        width = 5, 
#        height = 5)
```

